\batcave: Calling somatic mutations with a tumor- and site-specific prior

### 2019-10-08 NAR_DRAFT.tex
commit to merge

Many methods have been developed for calling somatic mutations from NGS data.
The earliest widely used somatic variant callers developed specifically for tumors, MuTect1~\cite{Cibulskis2013} and Varscan2~\cite{Koboldt2012}, used a combination of heuristic filtering and a model of sequencing errors to identify and score potential variants and set a threshold score designed to balance sensitivity and specificity.
Subsequent research gave rise to a number of alternate strategies, including haplotype-based calling \cite{Garrison2012},
joint genotype analysis (SomaticSniper \cite{Larson2012}, JointSNVMix2 \cite{Roth2012a}, Seurat \cite{Christoforides2013}, CaVEMan \cite{Jones2016}, and MuClone \cite{Dorri2019}), allele frequency-based analysis (Strelka \cite{Saunders2012}, LoFreq \cite{Wilm2012}, EBCall \cite{Shiraishi2013b}, deepSNV \cite{Gerstung2012}, LoLoPicker \cite{Carrot-Zhang2017}, and MuSE \cite{Fan2016}), and ensemble and deep learning methods (MutationSeq \cite{Ding2012}, BAYSIC \cite{Cantarel2014}, SomaticSeq \cite{Fang2015}, and SNooPer \cite{Spinella2016}).
These methods vary in their complexity and specific focus.
But they all implicitly or explicitly assume that the rate of mutation is uniform across the genome.

>>>>

Many methods have been developed for calling somatic mutations from NGS data.
The earliest widely used somatic variant callers developed specifically for tumors, MuTect1~\cite{Cibulskis2013} and Varscan2~\cite{Koboldt2012}, used a combination of heuristic filtering and a model of sequencing errors to identify and score potential variants and set a threshold score designed to balance sensitivity and specificity.
Subsequent research gave rise to a number of alternate strategies, including haplotype-based calling \cite{Garrison2012},
joint genotype analysis (SomaticSniper \cite{Larson2012}, JointSNVMix2 \cite{Roth2012a}, Seurat \cite{Christoforides2013}, CaVEMan \cite{Jones2016}, and MuClone \cite{Dorri2019}), allele frequency-based analysis (Strelka2 \cite{Kim2018}, LoFreq \cite{Wilm2012}, EBCall \cite{Shiraishi2013b}, deepSNV \cite{Gerstung2012}, LoLoPicker \cite{Carrot-Zhang2017}, and MuSE \cite{Fan2016}), and ensemble and deep learning methods (MutationSeq \cite{Ding2012}, BAYSIC \cite{Cantarel2014}, SomaticSeq \cite{Fang2015}, and SNooPer \cite{Spinella2016}).
These methods vary in their complexity and specific focus.
But they all implicitly or explicitly assume that the rate of mutation is uniform across the genome.
